A performance assessment of evolutionary algorithms in volunteer computing environments

### 2016-11-13 EvoPar2017/paper.Rnw
Reviewing text and eliminating comments

The architecture of the {\sf NodIO} system is divided in two tiers:\begin{enumerate}
\item A REST (representational state transfer) server that responds to
  HTTP requests made by the browsers. These HTTP requests use JSON for information 
  encoding, in this case chromosomes and related
  information sent by the clients and responses returned by the
  server.  The server has the capability to
  run a single experiment, storing the incoming chromosomes in a
  key-value cache that uses as key the chromosome string itself and is
  reset when the solution is found. This cache has a finite 
  size that erases the oldest chromosomes once it has been filled to % Also "it has been filled to its"
  capacity.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays the fitness and
  other graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client is doing. In
  the version of the server used in this paper, we implemented a
  policy by which the combination IP-fitness chromosome could only be
  inserted once into the cache. That means that if a particular IP
  sends the fitness 99 into the server, it will be accepted only
  once. The main intention of this policy was to avoid {\em
    overlapping} of contributions, with clients still running an {\em
    old} run of the experiment contributing after a new one had
  started. This did not completely avoid these contributions: if the
  clients of the {\em old} run increased their new fitness, it would
  still be allowed. 
  % As I remember we kept the run_number so IPs with old runs
  % did not interfere with newer ones. But you are saying: what if they are super slow 
  % and always PUT when the run is already over?, we want to receive 
  % those individuals? 
  % "That means that if a particular IP
  % sends the fitness 99 into the server, it will be accepted only
  % once."  
  % once per run?
% JJ - once per run or experiment. But we didn't keep that run_number
% to avoid interference, just to keep track
  %  but different individuals could have the same fitness, are we limiting the search? 
% We are actually checking for the same individual, not only the fitness
  %  maybe we could have multiple versions of the pool a la MVCC  :)
% Let's leave that for the future...- JJ
  % I think if we are just trying to limit overlapping of the experiments
  % then PUTs should send the run's seq_number and completly remove it.  
% The client does not keep track of the seq number. Maybe we should do
% just that, but sending back and forth the seq number implies
% refactoring. - JJ
\end{enumerate}

@@

graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 0.0520509 & 36 & 73 & 129.04545\\
\hline
reboot & 796214.0 & 9 & 6 & 646.0 & 67.0 & 0.0955121 & 45 & 168 & 75.88889\\
\hline
cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 0.1575092 & 55 & 225 & 43.28571\\
\hline
\end{tabular}
%% \end{tabular}
%% Initially generated with this code, 
% <<experiments, cache=FALSE,echo=FALSE>>=
%experiment.aggregates$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
%kable(experiment.aggregates)
%@ %def 
\end{center}
\end{table}
%
%
%
Since the point of using volunteer computing system is to cut costs,
all experiments were set up using the OpenShift
PaaS, which provides a free tier, making the whole experiment cost
equal to \$0.00. In fact, the NodIO system can be deployed to any
system as long as it can run node.js and has access to the filesystem, for
log files, if available%% . All these logs are available upon
%% request. Processed files are available in the same repository that
%% hosts this paper, as well as the scripts needed to process them.

>>>>

The architecture of the {\sf NodIO} system is divided in two tiers:\begin{enumerate}
\item A REST (representational state transfer) server that responds to
  HTTP requests made by the browsers. These HTTP requests use JSON for information 
  encoding, in this case chromosomes and related
  information sent by the clients and responses returned by the
  server.  The server has the capability to
  run a single experiment, storing the incoming chromosomes in a
  key-value cache that uses as key the chromosome string itself and is
  reset when the solution is found. This cache has a finite 
  size that erases the oldest chromosomes once it has been filled to % Also "it has been filled to its"
  capacity.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays the fitness and
  other graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client is doing. In
  the version of the server used in this paper, we implemented a
  policy by which the combination IP-fitness chromosome could only be
  inserted once into the cache. That means that if a particular IP
  sends the fitness 99 into the server, it will be accepted only
  once. The main intention of this policy was to avoid {\em
    overlapping} of contributions, with clients still running an {\em
    old} run of the experiment contributing after a new one had
  started. This did not completely avoid these contributions: if the
  clients of the {\em old} run increased their new fitness, it would
  still be allowed. 
\end{enumerate}

@@

graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 0.0520509 & 36 & 73 & 129.04545\\
\hline
reboot & 796214.0 & 9 & 6 & 646.0 & 67.0 & 0.0955121 & 45 & 168 & 75.88889\\
\hline
cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 0.1575092 & 55 & 225 & 43.28571\\
\hline
\end{tabular}
%% \end{tabular}
%% Initially generated with this code, 
% <<experiments, cache=FALSE,echo=FALSE>>=
%experiment.aggregates$Experiment <- c("base","graph1","graph2","reboot","cachereboot")
%kable(experiment.aggregates)
%@ %def 
\end{center}
\end{table}
%
%
%
Since the point of using volunteer computing system is to cut costs,
all experiments were set up using the OpenShift
PaaS, which provides a free tier, making the whole experiment cost
equal to nil. In fact, the NodIO system can be deployed to any
Platform or Infraestructure as a Service 
system as long as it can run node.js and can store logs on the same
place or in some other Log as a Service. Local storage can be used,
but it is not a requisite and, in fact, Heroku does not allow
persistent storage of files in its service. 
All these logs are available upon
request. Processed files are available in the same repository that
hosts this paper, as well as the scripts needed to process them.
